### 2014-01-29 inst/doc/pubs/one-simple-way.Rmd
still working on draft blog post

The second file provides the unprocessed data as a `.csv` file.  Because `csv` file formats vary

@@

- Standard table formats
- Standard cell formats

Rather than insist everyone use the ISO8601 format for dates, the user can declare whatever date format they prefer.

@@

R has good null values in its use of `data.frames`, so we already have
these.  Better yet, we write out our choice of null value in the metadata.
Reading in the data, we can explicitly pass this to `read.csv` rather than
just assume the null value is appropriate.  This is handled automatically
by `eml_read`.

@@

Context nodes serve the pupose described in this section.   
Machine-readable metadata,

@@

We have quality control on the metadata as well as the data.  While user
mistakes are still possible, automated validation of the metadata
provided and

>>>>

The second file provides the unprocessed data as a `.csv` file.
Because `csv` file formats can vary in the details of data orientation,
delimeters, and so forth, this information should also be provided as metadata
to ensure reliable and machine-readable intepretation.  The metadata generated 
using the EML package automatically adds this information based on R's handling
of csv files:

```xml
<objectName>data.csv</objectName>
<size>4984</size>
<characterEncoding>en_US.UTF-8</characterEncoding>
<dataFormat>
<textFormat>
  <numHeaderLines>1</numHeaderLines>
  <numFooterLines>0</numFooterLines>
  <recordDelimiter>\r\n</recordDelimiter>
  <attributeOrientation>column</attributeOrientation>
  <simpleDelimited>
    <fieldDelimiter>,</fieldDelimiter>
  </simpleDelimited>
</textFormat>
</dataFormat>
```

@@

- Standard table formats

- Standard cell formats

White _et al._ recommend that all users learn and adopt certain standard
formats for representing data, such as the ISO8601 standard for dates
(YYYY-MM-DD).  While we support this recommendation, this practice can be
not only burdensome to the data creator but also leaves any future user of
the data guessing or hoping that the format is one that they recognize.
Like most computer programs, R does not assume that any such pattern
is a date in ISO8601 format, but rather uses a Date type that permits
a variety of printing options.  Similarly, the `EML` package relies on
the metadata to state which cells contain dates and in what format those
dates are rendered.

Likewise, the package detects the appropriate metadata based on the R 
representation (such as factor, ordered factor, numeric, integer, character, or Date),
and writes the appropriate metadata for this format.

@@

R has good null values in its use of `data.frames`, and permits 
the user to identify other values for null or missing data.

Better yet, we write out our choice of null value in the metadata.

Reading in the data, we can explicitly pass this to `read.csv` rather than
just assume the null value is appropriate.  This is handled automatically
by `eml_read`.

@@

The detailed column and unit metadata serve the pupose described in this section of White _et al_.   
Machine-readable metadata,

@@

We have quality control on the metadata as well as the data.  While user
mistakes are still possible, automated validation of the metadata are
provided reduce the chance of errors.

The `EML` package can perform three levels of validation on EML files
written from or read into R: validating the XML, validating against the
EML Schema, and performing other quality control checks.  The `eml_validate`
function lets a user confirm that the metadata meets the EML standard,
helping to catch errors that could cause difficulty for another or user
or computer in intepreting the metadata:

```{r}
eml_validate("example.xml")
```

The `eml_validate` function is also run automatically any time a
user attempts to publish data to an external respository. MD5 checksums
of the data are also computed at this time to ensure that only complete 
and uncorrupted data is tranfered over the connection.

Several mechanisms are in place to evaluate the quality of the `EML`
package functions themselves.  In addition to the examples from the
documentation and the automated package checks provided for the R system,
the `EML` package includes a complete test suite using the testthat
package [Wickham 2011].  This helps ensures that the software works
as expected and maintains consistent behavior in any future versions
of the software package.

[Wickham 2011]: http://vita.had.co.nz/papers/testthat.html
